# Practical machine Learning: Project

## Synopsis

The goal of this project is to classify type of physical activity using Random Forest algorithm with data from wearable accelerometers. More information is available at:

https://class.coursera.org/predmachlearn-003

http://groupware.les.inf.puc-rio.br/har

## Data Cleaning

Preliminary data exploration showed that there are some number of empty data cells or cells filled with #DIV/0!". This fact is taken into account during reading of train and test data:

```{r message=FALSE}
library(randomForest)
train <- read.table("pml-training.csv",header=TRUE,sep=",", na.strings = c("NA","","#DIV/0!"))
test <- read.table("pml-testing.csv",header=TRUE,sep=",", na.strings = c("NA","","#DIV/0!"))
```

## Feature Selection

Further, data columns with at least one NA value were discarded:

```{r}
cols <- c()
for (col in colnames(train)) {
  if (!any(is.na(train[col]))) cols <- c(cols, col)
}
```

Finally, data columns which identify uniquely either the sibject ot the measurement moment were also discarded in order to avoid possible overfitting:

```{r warning=FALSE}
cols <- cols[! cols %in% c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","num_window")]
train <- train[cols]
```

## Model Building

Random Forest algorithm with default settings was used as a classifier:

```{r}
rf <- randomForest(classe ~ ., data=train)
predict(rf,newdata=test)
```

## Error estimation

One of the nice features of Random Forest algorithm is that error rate can be estimated from out-of-bag (OOB) samples:

```{r}
print(rf)
```

The error rate estimated this way is 0.3%

## Feature importances

```{r}
varImpPlot(rf)
```